### 2025-04-11 code/mesh_plotting_example.qmd
add mesh example

---
title: "plotly mesh plotting example"
author: "Rob Zielinski"
date: "04/07/2025"
---

```{r}
library(alphashape3d)
library(doParallel)
library(foreach)
library(Morpho)
library(plotly)
library(pme)
library(progress)
library(rgl)
library(Rfast)
library(tidyverse)

source("code/functions/fit_weighted_spline.R")
source("code/functions/print_SSD.R")
source("code/functions/fit_adni_additive_model.R")
source("code/functions/plot_id_pred.R")
source("code/functions/calculate_pme_reconstructions.R")

SSD_ratio_threshold <- 5
verbose <- TRUE

lhipp_surface <- read_csv("data_old/adni_fsl_lhipp_surface.csv")
adni_info <- read_csv("data_old/adni_fsl_info.csv")

lhipp_surface <- lhipp_surface |>
  mutate(
    date = decimal_date(date),
    Group = gsub("EMCI", "MCI", Group),
    Group = gsub("LMCI", "MCI", Group),
  )

lhipp_surface_centers <- lhipp_surface |>
  group_by(subid, date, scan_id) |>
  summarize(
    mean_x = mean(x),
    mean_y = mean(y),
    mean_z = mean(z),
    max_x = max(abs(x - mean_x)),
    max_y = max(abs(y - mean_y)),
    max_z = max(abs(z - mean_z))
  )

lhipp_bl <- lhipp_surface |>
  group_by(subid) |>
  arrange(date) |>
  summarize(
    date_bl = first(date),
    max_date = max(date)
  ) |>
  mutate(duration = max_date - date_bl)

lhipp_surface <- lhipp_surface |>
  full_join(lhipp_surface_centers, by = c("subid", "date", "scan_id")) |>
  full_join(lhipp_bl, by = "subid") |>
  mutate(
    x = (x - mean_x) / max_x,
    y = (y - mean_y) / max_y,
    z = (z - mean_z) / max_z,
    time_from_bl = date - date_bl,
    partition = ifelse(x > 0, 1, 2)
  )

lhipp_test <- lhipp_surface |>
  filter(image_id == "I13721") |>
  select(x, y, z) |>
  as.matrix()

ash <- ashape3d(
  lhipp_test,
  alpha = 0.25,
  pert = TRUE
)
test_mesh <- as.mesh3d(ash)
```

```{r}
plot_ly(
  x = test_mesh$vb[1, ],
  y = test_mesh$vb[2, ],
  z = test_mesh$vb[3, ],
  i = test_mesh$it[1, ] - 1,
  j = test_mesh$it[2, ] - 1,
  k = test_mesh$it[3, ] - 1,
  type = "mesh3d"
) |>
  layout(scene = list(aspectmode = "data"))
```

```{r}
plot_ly(
  x = lhipp_test[, 1],
  y = lhipp_test[, 2],
  z = lhipp_test[, 3],
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 3)
)
```

```{r}
data(nose)##load data
##warp a mesh onto another landmark configuration:
longnose.mesh <- tps3d(shortnose.mesh, shortnose.lm, longnose.lm,threads=1)
## Not run: 
mD <- meshDist(longnose.mesh, shortnose.mesh)
##now change the color ramp
render(mD,rampcolors = c("white","red"))

## End(Not run)
#use unsigned distances and a ramp from blue to red
#color distances < 0.01 green:
## Not run: 
meshDist(
  longnose.mesh, 
  shortnose.mesh, 
  rampcolors = c("blue", "red"),
  sign=FALSE, 
  tol=0.5
)
```